---
title: "CEE118X_HW4"
author: "Bill Li"
date: "10/15/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning=F, message = F) #warning messages and codes do not show
```

```{r}
library(tidyverse)
library(censusapi) #we are using census data
library(dplyr)
Sys.setenv(CENSUS_KEY="c8aa67e4086b4b5ce3a8717f59faa9a28f611dab")
```

```{r}
acs_vars_2019_5yr <-
  listCensusMetadata(
    name = "2019/acs/acs5",
    type = "variables"
  ) #Load our dictionary 
```

```{r}

#Income and race
#  datacensus.com to find household income b19001a=white only
bay_edu_race_2019<-getCensus(
  name = "acs/acs5", #five year survey
  vintage = "2019",
  region = "county: 001,013,041,055,075, 081, 085, 095, 097", #all SF counties
  regionin = "state:06", #defining the state
  vars = "group(C15002A)" # SEX BY EDUCATIONAL ATTAINMENT FOR THE POPULATION 25 YEARS AND OVER (WHITE ALONE)
)%>%
  select(ends_with("E") & -c(state,NAME)) %>% # I only want columns that ends with "E" also get rid of state and NAME column; We want bay area total income from white. -c removes the field of state and NAME. 
  summarize_all(sum) %>%# we sum up all the columns; It's gotta be numbers
  #we want to pivot the data into multiple rows and one column; make the dataset verticle
  pivot_longer(
    everything(),#pivot the entire data, everything will be transformed
    names_to = "name",
    values_to = "estimate" #household counts in this case. 
  )%>%
  # Use data dictionary acs_vars_2019_5yr to join names from dictionary with our data
  left_join(
    acs_vars_2019_5yr %>% 
      select(name,label) #we only want "name" and "label" in the dataset. 
  )%>%
  separate( #separate, we don't need Estimate!!Total:!!
    label, #it's called label from acs_vars_2019_5yr
    into = c(NA, NA, NA,"education"),
    sep = "!!"  #we re cognize the third one(after Total:!!) is income
  )%>%
  select(-name)%>%# at this point we don't want name anymore
  filter(!is.na(education))%>%
  group_by(education)%>%# got rid of empty field "NA"
  summarize(estimate = sum(estimate, na.rm=T))%>%
  mutate(
    race= "white alone" #assigning another column of white alone dataset. 
  )
#At this point we got just one race with income information 
```


```{r} 
#PIPELINE METHOD: (VERY GOOD METHOD)

race_categories <- c(
  "White alone",
  "Black or African American alone",
  "American Indian and Alaska Native alone",
  "Asian alone",
  "Native Hawaiian and Other Pacific Islander alone",
  "Some Other Race alone",
  "Two or more Races"
) #Create this for the ease of mutate below, so now race_categories[1] refers to white alone. 
#This is region code:  https://www.census.gov/prod/techdoc/cbp/cbp95/st-cnty.pdf 
bay_edu_race_2019<-
  1:7 %>%
  map_dfr(function(x){getCensus(
   name = "acs/acs5", #five year survey
  vintage = "2019",
  region = "county: 081",#081= San Mateo  #all SF counties: 001,013,041,055,075, 081, 085, 095, 097
  regionin = "state:06", #defining the state
  vars = paste0("group(C15002", LETTERS[x],")")# SEX BY EDUCATIONAL ATTAINMENT FOR THE POPULATION 25 YEARS AND OVER (WHITE ALONE)
)%>%
  select(ends_with("E") & -c(state,NAME)) %>% # I only want columns that ends with "E" also get rid of state and NAME column; We want bay area total income from white. -c removes the field of state and NAME. 
  summarize_all(sum) %>%# we sum up all the columns; It's gotta be numbers
  #we want to pivot the data into multiple rows and one column; make the dataset verticle
 pivot_longer(
    everything(),#pivot the entire data, everything will be transformed
    names_to = "name",
    values_to = "estimate" #household counts in this case. 
  )%>%
  # Use data dictionary acs_vars_2019_5yr to join names from dictionary with our data
  left_join(
    acs_vars_2019_5yr %>% 
      select(name,label) #we only want "name" and "label" in the dataset. 
  )%>%
  separate( #separate, we don't need Estimate!!Total:!!
    label, #it's called label from acs_vars_2019_5yr
    into = c(NA, NA, NA,"education"),
    sep = "!!"  #we re cognize the third one(after Total:!!) is income
  )%>%
  select(-name)%>%# at this point we don't want name anymore
  filter(!is.na(education))%>%
  group_by(education)%>%# got rid of empty field "NA"
  summarize(estimate = sum(estimate, na.rm=T))%>%
  mutate(
    race= race_categories[x] # use x to refer back to 1:7
    #you can put your code here
  )
  }
)
#automatically dataframe rbind(). Feed 1:7 into map_dft function and 

```

```{r}
#easy version
bay_edu_race_2019 %>% 
  group_by(education, race) %>% 
  summarize(estimate = sum(estimate)) %>% 
  ggplot() +
  geom_bar(
    aes(
      x = education %>% factor(levels = unique(bay_edu_race_2019$education)),
      y = estimate,
      fill = race %>% factor(levels = unique(bay_edu_race_2019$race))
    ),
    stat = "identity",
    position = "stack"
  ) +
  labs(
    x = "Education",
    y = "Individuals 25 years or older",
    title = "San Mateo County education attainment by race(stack) ",
    fill = "Race of Individual"
  ) +
  coord_flip()
```




```{r}
bay_edu_race_2019 %>% 
  group_by(education, race) %>% 
  summarize(estimate = sum(estimate)) %>% 
  ggplot() +
  geom_bar(
    aes(
      x = education %>% factor(levels = rev(unique(bay_edu_race_2019$education))),
      y = estimate,
      fill = race %>% factor(levels = rev(unique(bay_edu_race_2019$race)))
    ),
    stat = "identity",
    position = "fill"
  ) +
  labs(
    x = "Education",
    y = "Proportion of Individuals 25 years or older",
    title = "San Mateo County education attainment by race(fill)",
    fill = "Race of Individual"
  ) +
  coord_flip() +
  theme(
    legend.position = "bottom",
    legend.direction = "vertical"
  )  +
  guides(
    fill = guide_legend(
      reverse = T
    )
  )
```

“Is there disproportionate education attainment by race in the county?”
Answer: Based on the data from San Mateo County, there is a disproportionate education attainment by race for individuals 25 years or older. In white race population, there is a significant more proportion of individuals that have some college or bachelor's degree than lss than high school diploma. When you compare white alone to black or African American alone, there is a disproportionate education attainment. The proportion of black individuals holding bachelor's degree or higher is far less than high school graduate. The situation is the opposite in white individuals. American Indian and Alaska Native alone does not face disproportionate education attainment. In Asian alone population, there is a greater proportion of individuals that have bachelor's degree or higher than the three other educational categories. 