---
title: "CEE118X_HW6"
author: "Bill Li"
date: "10/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning=F, message = F) #warning messages and codes do not show
```

```{r}
library(censusapi)
library(tidyverse)
library(tigris)
library(sf)
library(leaflet)

Sys.setenv(CENSUS_KEY="c8aa67e4086b4b5ce3a8717f59faa9a28f611dab")
```

```{r}
Sys.setenv(CENSUS_KEY="c8aa67e4086b4b5ce3a8717f59faa9a28f611dab")
pums_2019_1yr <- getCensus(
  name = "acs/acs1/pums",
  vintage = 2019,
  region = "public use microdata area:*", 
  regionin = "state:06",
  vars = c(
    "SERIALNO",
    "SPORDER",
    "PWGTP",
    "WGTP",
    "AGEP",
    "YBL", 
    "BLD", 
    "TEN", 
    "MV", 
    "HINCP", 
    "AGEP"
  )
)
```

```{r}
ca_pumas <-
  pumas("CA", cb = T, progress_bar = F)

ca_pumas <-
  pumas("CA", cb = T, progress_bar = F)

sf_boundary <-
  counties("CA", cb = T, progress_bar = F) %>%
  filter(NAME == "San Francisco") #we only want SF

sf_pumas <-
  ca_pumas %>% 
  st_centroid() %>% 
  .[sf_boundary, ] %>% 
  st_drop_geometry() %>% 
  left_join(ca_pumas %>% select(GEOID10)) %>% 
  st_as_sf()
```
```{r}
sf_pums<-pums_2019_1yr %>% 
  mutate(
    PUMA = str_pad(public_use_microdata_area,5,"left","0")
  ) %>% 
  filter(PUMA %in% sf_pumas$PUMACE10) 
```

```{r}
sf_pums_clean<- sf_pums%>%
  mutate(
    YBL= as.numeric(YBL),
    AGEP= as.numeric(AGEP)
  )%>%
  filter(YBL %in% as.character(1:3))%>% #built before 1960
#or  YBL = as.numeric(YBL) %>% filter(YBL %in%  1:3) which is another method
#how to groupby and summarize with families of less than 6 year old. For the 
#youngest person in the household, are they 6?
  arrange(AGEP)%>% #arrange by age which makes the smallest age the first
  group_by(SERIALNO)%>% #serial number in the sf_pums
  summarize_all(first)%>%
  mutate(
    leadrisk= ifelse(
      as.numeric(HINCP)<90000 & as.numeric(AGEP)<6,
      1,
      0
    )
  ) #complete leadrisk
```

```{r}
#factor transformation
sf_p<- sf_pums_clean%>%
  mutate(
    Unit= BLD%>%
      factor(
        levels=sf_pums_clean$BLD %>%
          unique()%>%
          as.numeric()%>%
          sort()
      ))%>%
  mutate(
    Tenure= TEN%>%
      factor(
        levels=sf_pums_clean$TEN %>%
          unique()%>%
          as.numeric()%>%
          sort()
      ))%>%
  mutate(
    Move= MV%>%
      factor(
        levels=sf_pums_clean$MV %>%
          unique()%>%
          as.numeric()%>%
          sort()
      ))%>%
  mutate(
    Microdata= PUMA%>%
      factor(
        levels=sf_pums_clean$PUMA %>%
          unique()%>%
          sort()
      )) #complete factor transformation
```


```{r}
#logit model
logit_model<- glm(
  leadrisk ~ Unit+Tenure+Move+Microdata,
  family= quasibinomial(),
  data= sf_p
)

summary(logit_model) #logit model complete

```

```{r}
predict(logit_model, data.frame(sample_n(sf_p,1)), type="response")
  
```

```{r}
nrow(sf_p)
#there are 2264 rows of data
```

```{r}
Prediction<-predict(logit_model, data.frame(sample_n(sf_p,2264)), type="response")
#
sf_final<- sf_p%>%
  cbind(Prediction)

sf_2by2<- sf_final%>%
  mutate(
    leadrisk=ifelse(
      leadrisk==1,
      "Positive",
      "Negative"
    )
  )%>%
  mutate(
    Prediction=ifelse(
      Prediction<0.1,
      "True",
      "False"
)
)%>%
  group_by(leadrisk,Prediction)%>%
  summarize(WGTP=sum(as.numeric(WGTP), na.rm=T))
```
Based on threshold score along, they will mail out (221569+4258) 225,827 postcards based on threshold score. 4258/22886 or 18.6 percent of the postcards will go to low income household with children. 4258/4409 or 96.6 percent of all at-risk children are likely to be reached. To make sure all postcards will go to at-risk children, we can consider adding a new category in the survey to reflect family's at-risk situation. A new category can make the prediction more stringent and reduce the number of false-positive counts. 

```{r}
sum(sf_2by2$WGTP)
```






#
sample<- sample(
  c(TRUE, FALSE),
  nrow(sf_p),
  replace=T,
  prob=c(0.8,0.2)
)
train<- sf_p[sample,]
test<- sf_p[!sample,]




#
train_design<- svrepdesign(
  data=train,
  type="ACS",
  repweights= 



sum<-test%>%
  mutate(
    leadrisk= ifelse(
      leadrisk==1,
      "Yes",
      "No"
    )
  )%>%
  pull(leadrisk)%>%
  table(Prediction>0.5)





