---
title: "CEE118X_Session5"
author: "Bill Li"
date: "10/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning=F, message = F) #warning messages and codes do not show
```

Use PUMS data to measure individual response(access to AC...)


```{r}
library(tidyverse)
library(censusapi) #we are using census data
Sys.setenv(CENSUS_KEY="c8aa67e4086b4b5ce3a8717f59faa9a28f611dab")
```


FOR WHITE ONLY Code
#Income and race
#  datacensus.com to find household income b19001a=white only
bay_income_race_2019<-getCensus(
  name = "acs/acs5", #five year survey
  vintage = "2019",
  region = "county: 001,013,041,055,075, 081, 085, 095, 097", #all SF counties
  regionin = "state:06", #defining the state
  vars = "group(B19001A)" #income by white along household(represent by "A")
)%>%
  select(ends_with("E") & -c(state,NAME)) %>% # I only want columns that ends with "E" also get rid of state and NAME column; We want bay area total income from white. -c removes the field of state and NAME. 
  summarize_all(sum) %>%# we sum up all the columns; It's gotta be numbers
  #we want to pivot the data into multiple rows and one column; make the dataset verticle
  pivot_longer(
    everything(),#pivot the entire data, everything will be transformed
    names_to = "name",
    values_to = "estimate" #household counts in this case. 
  )%>%
  # Use data dictionary acs_vars_2019_5yr to join names from dictionary with our data
  left_join(
    acs_vars_2019_5yr %>% 
      select(name,label) #we only want name and label in the dataset. 
  )%>%
  separate( #separate, we don't need Estimate!!Total:!!
    label, #it's called label from acs_vars_2019_5yr
    into = c(NA, NA, "income"),
    sep = "!!"  #we recognize the third one(after Total:!!) is income
  )%>%
  select(-name)%>% # at this point we don't want name anymore
  # The first row is sum of the total households
  filter(!is.na(income))%>% # remove the empty field in "income" column. NEED TO ADD ! to remove empty field!!!
  mutate(
    race= "white alone" #assigning another column of white alone dataset. 
  )
#At this point we got just one race with income information 


USING For Loop:
for (x in LETTERS[1:7]){ #R supplies Letter[] , Letter[1] represents "A"
  LETTERS[x] #Will refer back to LETTERS[1:7]
  
  #Eventually,
  rbind(full_dataset) which combines all data
  

```{r} 
#PIPELINE METHOD: (VERY GOOD METHOD)

race_categories <- c(
  "White alone",
  "Black or African American alone",
  "American Indian and Alaska Native alone",
  "Asian alone",
  "Native Hawaiian and Other Pacific Islander alone",
  "Some Other Race alone",
  "Two or more Races"
) #Create this for the ease of mutate below, so now race_categories[1] refers to white alone. 

bay_income_race_2019<-
  1:7 %>%
  map_dfr(function(x){
    getCensus(
  name = "acs/acs5", #five year survey
  vintage = "2019",
  region = "county: 001,013,041,055,075, 081, 085, 095, 097", #all SF counties
  regionin = "state:06", #defining the state
  vars = paste0("group(B19001",LETTERS[x],") ")#income by white along household(represent by "A")
)%>%
  select(ends_with("E") & -c(state,NAME)) %>% # I only want columns that ends with "E" also get rid of state and NAME column; We want bay area total income from white. -c removes the field of state and NAME. 
  summarize_all(sum) %>%# we sum up all the columns; It's gotta be numbers
  #we want to pivot the data into multiple rows and one column; make the dataset verticle
  pivot_longer(
    everything(),#pivot the entire data, everything will be transformed
    names_to = "name",
    values_to = "estimate" #household counts in this case. 
  )%>%
  # Use data dictionary acs_vars_2019_5yr to join names from dictionary with our data
  left_join(
    acs_vars_2019_5yr %>% 
      select(name,label) #we only want name and label in the dataset. 
  )%>%
  separate( #separate, we don't need Estimate!!Total:!!
    label, #it's called label from acs_vars_2019_5yr
    into = c(NA, NA, "income"),
    sep = "!!"  #we recognize the third one(after Total:!!) is income
  )%>%
  select(-name)%>% # at this point we don't want name anymore
  # The first row is sum of the total households
  filter(!is.na(income))%>% # remove the empty field in "income" column. NEED TO ADD ! to remove empty field!!!
  mutate(
    race= race_categories[x] # use x to refer back to 1:7
    #you can put your code here
  )
  }
)
#automatically dataframe rbind(). Feed 1:7 into map_dft function and 

```

```{r}

```

