---
title: "CEE118X_HW9"
runtime: shiny
output: 
  flexdashboard::flex_dashboard:
    vertical_layout: fill
---
```{r}
#load library shiny 
#```{r}
#bay_zips.rds creation 

#run this again!!!!!
library(tigris)
library(tidyverse)
#ca_zips<-zctas("CA")
#bay_county_names<- c("Aldameda","Contra Costa", "Marin","Napa","San Francisco","San Mateo", "Santa Clara", "Solano","Sonoma")
#bay_counties<- counties("CA")%>% filter(NAME%in% bay_county_names)
#bay_zips<- ca_zips[bay_counties,]
#saveRDS(bay_zips, "bay_zips.rds")
#```
```

```{r global, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(leaflet)
library(sf)
library(plotly)

pge_data_raw <- readRDS("pge_data_raw.rds")

bay_zips <- readRDS("bay_zips.rds")

pge_data <-
  pge_data_raw %>% 
  filter(
    CUSTOMERCLASS %in% c(
      "Elec- Commercial",
      "Elec- Residential",
      "Gas- Commercial",
      "Gas- Residential"
    )
  ) %>% 
  group_by(
    MONTH, 
    YEAR, 
    CUSTOMERCLASS
  ) %>% 
  summarize(
    TOTALKBTU = sum(TOTALKBTU, na.rm = T)
  ) %>% 
  mutate(
    DATE = 
      paste(
        YEAR,
        MONTH, 
        "01",
        sep="-"
      ) %>% as.Date()
  )
```

Inputs {.sidebar}
-------------------------------------

```{r}
library(shiny)
selectInput(
  inputId = "year",
  label = "Year: ",
  choices = 2017:2020,
  selected = 2020
)
```




Column
-------------------------------------

### Monthly consumption


```{r}
#make plot reactive with year change
#actual output 
plotlyOutput("plot")
```



```{r, context = "server"}
observeEvent(
  input$year, {
  chart <- pge_data %>% 
  filter(
    CUSTOMERCLASS %in% c(
      "Elec- Residential"
    ),
    YEAR == input$year  #for multiple years 
  ) %>% 
  ggplot(
    aes(
      x = MONTH,
      y = TOTALKBTU/1e9
    )
  ) +
  geom_line(
    aes(
      color = YEAR %>% factor()
    )
  ) +
  scale_x_discrete(
    limits = c(
      "Jan",
      "Feb",
      "Mar",
      "Apr",
      "May",
      "Jun",
      "Jul",
      "Aug",
      "Sep",
      "Oct",
      "Nov",
      "Dec"
    )
  ) +
  labs(
    x = "",
    y = "Total kBTUs (billions)",
    title = paste0("Residential Energy Consumption in the Bay Area, ",input$year),
    color = "Year"
  ) + 
  theme(legend.position = "none")

  output$plot<- renderPlotly({
    chart %>% 
  ggplotly() %>% 
  config(displayModeBar = F)
  })
  
})
#make it behind the scene with context= "server". 
#now when you use the dropdown menu of the year, it will change the plot 
```

Column
-------------------------------------

### ZCTA consumption

```{r}
leafletOutput("map") # do the same thing with render() similar to above
```


```{r, contex = "server"}
observeEvent(input$year, {
  pge_20_res_elec <-
  pge_data_raw %>% 
  filter(
    CUSTOMERCLASS == "Elec- Residential",
    YEAR == input$year
  ) %>% 
  mutate(
    ZIPCODE = ZIPCODE %>% as.character()
  ) %>% 
  group_by(ZIPCODE) %>% 
  summarize(
    TOTALKBTU = sum(TOTALKBTU, na.rm = T)
  ) %>% 
  right_join(
    bay_zips %>% select(GEOID10),
    by = c("ZIPCODE" = "GEOID10")
  ) %>% 
  st_as_sf() %>% 
  st_transform(4326)

res_pal <- colorNumeric(
  palette = "Reds",
  domain = 
    pge_20_res_elec$TOTALKBTU
)

output$map <- renderLeaflet({
  leaflet() %>% 
  addProviderTiles(provider = providers$CartoDB.Positron) %>% 
  addPolygons(
    data = pge_20_res_elec,
    fillColor = ~res_pal(TOTALKBTU),
    color = "white",
    opacity = 0.5,
    fillOpacity = 0.5,
    weight = 1,
    label = ~paste0(
      round(TOTALKBTU), 
      " kBTU total in ",
      ZIPCODE
    ),
    highlightOptions = highlightOptions(
      weight = 2,
      opacity = 1
    )
  ) %>% 
  addLegend(
    data = pge_20_res_elec,
    pal = res_pal,
    values = ~TOTALKBTU,
    title = paste0("Total Residential<br>Electricity (kBTU),", input$year)
  )
})


  
})


```