---
title: "CEE118X_HW1"
author: "Bill Li"
date: "9/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning=F, message = F)
```

```{r}
library(tidyverse)
```

```{r}
years<-2017:2020
quarters<-1:4
type1<-"Electric"
pge_17to20_elec <- NULL #reserve the filename

for(quarter in quarters)
  for(year in years){ #loop with years and quarters
  filename<- paste0(
    "PGE_",
    year,
    "_Q",
    quarter,
    "_",
    type1,
    "UsageByZip.csv"
  ) 
  print(filename)
  temp<-read_csv(filename) #made all 4 quarters of PGE data into one csv
  pge_17to20_elec<-rbind(pge_17to20_elec, temp) #combine 4 csv files into one in global environment 
  write_csv(pge_17to20_elec, "pge_17to20_elec.csv") #save into csv format, not necessary
}
```
*Just created pge_17to20_ele.csv so from 2017 to 2020 electricity data.

```{r}
years<-2017:2020
quarters<-1:4
type<-"Gas"
pge_17to20_gas <- NULL #reserve the filename

for(quarter in quarters)
  for(year in years){ #loop with years and quarters
  filename1<- paste0(
    "PGE_",
    year,
    "_Q",
    quarter,
    "_",
    type,
    "UsageByZip.csv"
  )
  print(filename1)
  temp1<-read_csv(filename1) #made all 4 quarters of PGE data into one csv
  pge_17to20_gas<-rbind(pge_17to20_gas, temp1) #combine 4 csv files into one in global environment 
  write_csv(pge_17to20_gas, "pge_17to20_gas.csv") #save into csv format, not necessary
}
```
*Just created pge_17to20_gas.csv so from 2017 to 2020 gas data.

```{r}
library(dplyr)
pge_elec_final<- select(
  pge_17to20_elec, 
  -c(ZIPCODE,COMBINED,AVERAGEKWH) #use select() to get rid of columns
)%>%
  filter(   #use filter() to get rid of subcategories
    CUSTOMERCLASS %in%
      c(
        "Elec- Commercial", "Elec- Residential"
      )
    )%>%
  group_by(YEAR, MONTH, CUSTOMERCLASS)%>% #Use group_by() to avoid system getting rid of month and customerclass in summarize() function
  summarize(
  TOTALCUSTOMERS=
    sum(TOTALCUSTOMERS,
        na.rm = T),
  TOTALKWH=
    sum(TOTALKWH,
        na.rm = T) )%>%
  mutate(
    TOTALKBTU=
      TOTALKWH*3.412
  )%>%
  mutate(                   #Use paste() to combine year and month data
    DATE=
      paste(
        YEAR,
        MONTH,
        "01",
        sep="-"
      )%>% as.Date()
  )
  
  
```

```{r}
pge_gas_final<- select(
  pge_17to20_gas, 
  -c(ZIPCODE,COMBINED,AVERAGETHM)
)%>%
  filter(
    CUSTOMERCLASS %in%
      c(
        "Gas- Commercial", "Gas- Residential"
      ))%>%
  group_by(YEAR, MONTH, CUSTOMERCLASS)%>%
  summarize(
    TOTALCUSTOMERS= 
      sum(TOTALCUSTOMERS,
          na.rm = T),
    TOTALTHM=
      sum(TOTALTHM,
          na.rm = T)) %>%
      mutate(
        TOTALKBTU=
          TOTALTHM*100
      )%>%
  mutate(
    DATE=
      paste(
        YEAR,
        MONTH,
        "01",
        sep="-"
      )%>% as.Date()
  )
  
  
```
```{r}
pge_elec_res<- select(pge_elec_final,
                     -c(TOTALKWH))%>%
  filter(CUSTOMERCLASS %in%
                       c("Elec- Residential")) #Residential electricity data
pge_elec_com<- select(pge_elec_final,
                     -c(TOTALKWH))%>%
  filter(CUSTOMERCLASS %in%
                       c("Elec- Commercial")) #Commercial electricity data
```

```{r}
pge_gas_res<- filter(pge_gas_final, 
                     CUSTOMERCLASS %in%
                       c("Gas- Residential")) %>% #Residential gas data
  select(-c(TOTALTHM))
pge_gas_com<- filter(pge_gas_final, 
                     CUSTOMERCLASS %in%
                       c("Gas- Commercial")) %>% #Commercial gas data
  select(-c(TOTALTHM))
```

```{r}
pge_res<- rbind(pge_elec_res, pge_gas_res) #use arrange here to organize data
pge_com<- rbind(pge_elec_com, pge_gas_com)
```

```{r}
pge_res_chart<- pge_res %>% #only display electricity data
  ggplot()+
  geom_bar(
    aes(
      x=DATE %>% factor(),
      y=TOTALKBTU,
      fill= CUSTOMERCLASS 
    ),
    stat="identity",
    position="stack"
  )+
  labs(
    x="Month",
    y="Total KBTU",
    title= "PG&E Residential Gas and Electricity Usage, 2017-2020",
    fill="Electricity Type"
  )+
  scale_x_discrete(guide = guide_axis(n.dodge=7))
pge_res_chart 

```

```{r}
pge_com_chart<- pge_com %>% #only display electricity data
  ggplot()+
  geom_bar(
    aes(
      x=DATE %>% factor(),
      y=TOTALKBTU,
      fill= CUSTOMERCLASS 
    ),
    stat="identity",
    position="stack"
  )+
  labs(
    x="Month",
    y="Total KBTU",
    title= "PG&E Commercial Gas and Electricity Usage, 2017-2020",
    fill="Electricity Type"
  )+
  scale_x_discrete(guide = guide_axis(n.dodge=7))
pge_com_chart 
```

```{r}

pge_res_elec_compare<-pge_res%>%
  filter(CUSTOMERCLASS%in%
           c("Elec- Residential")) %>%
  ggplot()+   #Always define x and y axis in geom_line()
  geom_line(aes(
    x= MONTH,
    y= TOTALKBTU,
    colour=factor(YEAR)
  ),
  stat="identity") +
  labs(
    x="Month",
    y="Total KBTU",
    Title="PG&E Residential Electricity Usage 2017-2020"
  )
pge_res_elec_compare



#pge_com_compare
```

Ans:Based on the line chart above, residential electricity usage went up in 2020 from August to December comparing to the data from 2018 and 2019. There seems to be a sharp increase in residential electricity usage in 2017 between August and October, but it is more likely to be an anomoly. 

```{r}
pge_res_gas_compare<- pge_res%>%
  filter(CUSTOMERCLASS%in%
           c("Gas- Residential")) %>%
  ggplot()+   #Always define x and y axis in geom_line()
  geom_line(aes(
    x= MONTH,
    y= TOTALKBTU,
    colour=factor(YEAR)
  ),
  stat="identity") +
  labs(
    x="Month",
    y="Total KBTU",
    Title="PG&E Residential Gas Usage 2017-2020"
  )
pge_res_gas_compare
  
```

Ans: Based on the line chart above, residential gas usage in 2020 is about the same as the rest of the years with a slight increase from late October to December. There is a sharp increase in gas usage between August and October in 2017, but that is more likely to be an anomaly in the PG&E data. 


```{r}
pge_covid_res<- pge_res%>%
  filter(YEAR%in%
           c(2020))
```

```{r}
pge_covid_com<- pge_com%>%
  filter(YEAR%in%
           c(2020))
```

```{r}
pge_res_covid_chart<- pge_covid_res %>% #only display electricity data
  ggplot()+
  geom_bar(
    aes(
      x=MONTH %>% factor(),
      y=TOTALKBTU,
      fill= CUSTOMERCLASS 
    ),
    stat="identity",
    position="stack"
  )+
  labs(
    x="Month",
    y="Total KBTU",
    title= "PG&E Residential Gas and Electricity Usage 2020",
    fill="Electricity Type"
  )
pge_res_covid_chart 
```

```{r}
pge_com_covid_chart<- pge_covid_com %>% #only display electricity data
  ggplot()+
  geom_bar(
    aes(
      x=MONTH %>% factor(),
      y=TOTALKBTU,
      fill= CUSTOMERCLASS 
    ),
    stat="identity",
    position="stack"
  )+
  labs(
    x="Month",
    y="Total KBTU",
    title= "PG&E Commercial Gas and Electricity Usage 2020",
    fill="Electricity Type"
  )
pge_com_covid_chart 
```

```{r}
par(mfrow=c(1,2))
pge_res_chart
pge_res_covid_chart
```

```{r}
par(mfrow=c(1,2))
pge_com_chart
pge_com_covid_chart
```

*ANS: For both residential and commercial electricity and gas usage, the general trend is similar for pre-pandemic and covid era. Both Commercial and Residential use more gas in the winter and more electricity in the summer.
In the 2017-2020 data, I did not distinguish the various years in my data. As a result, variations in certain months of years can be mask by the chart. We assume that monthly data are similar among these years in this case. 
